---
title: "Bayesian analysis for a semiparametric meta-analysis model using a Dirichlet Process Mixture of Normals prior"
author: 
  - name: Joon-Ho Lee (joonho@berkeley.edu)
date: "February 28, 2020"
output:
  html_document: 
    css: styles.css
    fig_caption: yes
    highlight: haddock
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
  tufte::tufte_html:
    number_sections: yes
    toc: true
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '3'
bibliography: bibliography.bib
---

```{r basic_setup, include=FALSE}
### Set working directory
setwd("~/meta-analysis/01_DPmeta")

### Set RMarkdown options
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)

### Call libraries
library(tidyverse)
library(cowplot)
library(DPpackage)

# ### Set Stan options
# # options(mc.cores = parallel::detectCores())
# rstan_options(auto_write = TRUE)
# Sys.setenv(LOCAL_CPPFLAGS = '-march=native')

### Theme settings
theme_preset <- 
  theme_bw() + 
  theme(panel.background = element_blank(),
        panel.grid = element_blank(), 
        legend.position = "bottom", 
        legend.direction = "horizontal", 
        legend.title = element_blank())
```


# `DPmeta`

This function generates a posterior density sample for a semiparametric linear mixed effects metaanalysis model using a Dirichlet process or a Mixture of Dirichlet process prior for the distribution of the random effects. 

```{r}
##################################################################
# Data on the effectiveness of silver sulfadiazine coating
# on venous catheters for preventing bacterial colonisation of
# the catheter and bloodstream infection.
# Veenstra D et al (1998) "Efficacy of Antiseptic Impregnated
# Central Venous Catheters in Preventing Nosocomial Infections:
# A Meta-analysis" JAMA 281:261-267.
#
# Note that -Inf and Inf have been replaced by NA.
##################################################################


studies <- c("Tennenberg","Maki","vanHeerden", 
             "Hannan","Bach(a)","Bach(b)", 
             "Heard","Collins","Ciresi","Ramsay", 
             "Trazzera","George")

logOR <- c(-1.5187189,-0.7136877,-1.3217558,-0.1910552,NA,
           -2.2005195,-0.5057461,-2.3538784,-0.3643810, 
           -0.5371429,-0.7608058,-2.1400662)

varlogOR <- c(0.4157541,0.2632550,0.6739189,0.3727788,NA,
              0.7623470,0.2306169,0.7477891,0.3645463,0.2291839, 
              0.3561542,0.5190489)^2

names(logOR) <- studies
names(varlogOR) <- studies

y <- cbind(logOR,varlogOR)

colnames(y) <- c("logOR","varlogOR")

prior<-list(alpha=1, tau1=20,tau2=10,mub=0,Sb=100)

# Initial state
state <- NULL

# MCMC parameters
nburn<-20000
nsave<-10000
nskip<-20
ndisplay<-100

mcmc <- list(nburn=nburn, 
             nsave=nsave,
             nskip=nskip,
             ndisplay=ndisplay)

# Fit the model: First run
fit1<-DPmeta(formula=y~1,prior=prior,mcmc=mcmc,state=state,status=TRUE)
fit1
# Summary with HPD and Credibility intervals
summary(fit1)
summary(fit1,hpd=FALSE)

# Plot model parameters (to see the plots gradually set ask=TRUE)
plot(fit1,ask=FALSE)
plot(fit1,ask=FALSE,nfigr=2,nfigc=2)

# Prior information
```


# `DPMmeta`

This function generates a posterior density sample for a semiparametric linear mixed effects metaanalysis model using a Dirichlet Process Mixture of Normals prior for the distribution of the random effects. 

```{r}
##############################################################    
# Data on the effectiveness of silver sulfadiazine coating
# on venous catheters for preventing bacterial colonisation of 
# the catheter and bloodstream infection. 
# Veenstra D et al (1998) "Efficacy of Antiseptic Impregnated 
# Central Venous Catheters in Preventing Nosocomial Infections: 
# A Meta-analysis" JAMA 281:261-267. 
#
# Note that -Inf and Inf have been replaced by NA.
##################################################################    
    
studies <- c("Tennenberg","Maki","vanHeerden",
              "Hannan","Bach(a)","Bach(b)",
              "Heard","Collins","Ciresi","Ramsay",
              "Trazzera","George")    

logOR <- c(-1.5187189,-0.7136877,-1.3217558,-0.1910552,
            NA,-2.2005195,-0.5057461,-2.3538784,-0.3643810,
            -0.5371429,-0.7608058,-2.1400662)
 
varlogOR <- c(0.4157541,0.2632550,0.6739189,0.3727788,NA,
              0.7623470,0.2306169,0.7477891,0.3645463,0.2291839,
              0.3561542,0.5190489)^2

names(logOR) <- studies
names(varlogOR) <- studies
y <- cbind(logOR,varlogOR)
colnames(y) <- c("logOR","varlogOR")

# Prior information

prior<-list(alpha=1,
            tau01=20,
            tau02=10,
            tau11=20,
            tau12=10,
            mb=0,
            Sb=100)

# Initial state
state <- NULL


# MCMC parameters

nburn<-20000
nsave<-10000
nskip<-20
ndisplay<-100
mcmc <- list(nburn=nburn,
             nsave=nsave,
             nskip=nskip,
             ndisplay=ndisplay)

# Fit the model: First run

fit1<-DPMmeta(formula=y~1,prior=prior,mcmc=mcmc,
              state=state,status=TRUE)
fit1

# Summary with HPD and Credibility intervals
summary(fit1)
summary(fit1,hpd=FALSE)

# Plot model parameters (to see the plots gradually set ask=TRUE)
plot(fit1,ask=FALSE)
plot(fit1,ask=FALSE,nfigr=2,nfigc=2)	
```


# `PTmeta`

This function generates a posterior density sample for a semiparametric linear mixed effects meta-analysis model using a Polya Tree or a Mixture of Polya Trees prior for the distribution of the random effects.


```{r}
##############################################################    
# Data on the effectiveness of silver sulfadiazine coating
# on venous catheters for preventing bacterial colonisation of 
# the catheter and bloodstream infection. 
# Veenstra D et al (1998) "Efficacy of Antiseptic Impregnated 
# Central Venous Catheters in Preventing Nosocomial Infections: 
# A Meta-analysis" JAMA 281:261-267. 
#
# Note that -Inf and Inf have been replaced by NA.
##################################################################    

studies <- c("Tennenberg","Maki","vanHeerden",
              "Hannan","Bach(a)","Bach(b)",
              "Heard","Collins","Ciresi","Ramsay",
              "Trazzera","George")    

logOR <- c(-1.5187189,-0.7136877,-1.3217558,-0.1910552,
            NA,-2.2005195,-0.5057461,-2.3538784,-0.3643810,
            -0.5371429,-0.7608058,-2.1400662)
 
varlogOR <- c(0.4157541,0.2632550,0.6739189,0.3727788,NA,
              0.7623470,0.2306169,0.7477891,0.3645463,0.2291839,
              0.3561542,0.5190489)^2

names(logOR) <- studies
names(varlogOR) <- studies
y <- cbind(logOR,varlogOR)
colnames(y) <- c("logOR","varlogOR")


# Initial state
state <- NULL

# MCMC parameters

nburn<-20000
nsave<-10000
nskip<-20
ndisplay<-100
mcmc <- list(nburn=nburn,
             nsave=nsave,
             nskip=nskip,
             ndisplay=ndisplay)

# Prior information 1: non-median zero PT

prior1<-list(alpha=1,
             tau1=20,
             tau2=10,
             mub=0,
             Sb=100,
             M=4)

# Prior information 2: median zero PT

prior2<-list(alpha=1,
             tau1=20,
             tau2=10,
             mub=0,
             Sb=100,
             M=4,
             frstlprob=TRUE,
             Sbeta0=diag(1000,1),
             beta0=rep(0,1))
             

# Fitting the models

fit1<-PTmeta(formula=y~1,prior=prior1,mcmc=mcmc,
             state=state,status=TRUE)
fit1


fit2<-PTmeta(formula=y~1,prior=prior2,mcmc=mcmc,
             state=state,status=TRUE)
fit2

# Summary with HPD and Credibility intervals
summary(fit1)
summary(fit1,hpd=FALSE)

summary(fit2)
summary(fit2,hpd=FALSE)

# Plot model parameters (to see the plots gradually set ask=TRUE)
plot(fit1,ask=FALSE)
plot(fit1,ask=FALSE,nfigr=2,nfigc=2)	

plot(fit2,ask=FALSE)
plot(fit2,ask=FALSE,nfigr=2,nfigc=2)
```



# References



